---
title: "Anaylsis of AMD"
author: "Niall Graham"
date: "09 October 2017"
output: html_document
---

Loading and preprocessing the data

Set Working Directory 

```{r}
echo = TRUE
activity <- NULL
activity <- read.csv("activity.csv", header = T, sep = ",")
```

Variables created during by this code are set to NULL
```{r}
echo = TRUE
df_summary <- NULL
su2 <- NULL
su <- NULL
mn_int <- NULL
activity2 <- NULL
mean_su2 <- NULL
median_su2 <- NULL
activity2_weekend <- NULL
activity2_weekday <- NULL
mean_activity2_weekday <- NULL
mean_activity2_weekend <- NULL
```

1. Total Number of Steps Taken Per Day
```{r}
echo = TRUE
su <- tapply(activity$steps, activity$date, sum, na.rm=T)
echo = TRUE
hist(su, xlab = "sum of steps per day", main = "histogram of steps per day",col="red")
```


Mean Total Number of Steps 

```{r}
echo = TRUE
mean_su <- round(mean(su))
median_su <- round(median(su))
print(c("The mean is",mean_su))
```

Median Total Number of Steps

```{r}
print(c("The median is",median_su))
```

2. Average Daily Activity Pattern

```{r}
echo = TRUE
mn_int <- tapply(activity$steps, activity$interval, mean, na.rm=T)
plot(mn_int ~ unique(activity$interval), type="l", xlab = "5-min interval", col="blue")
```

2.b. 5- Minute Intervals

```{r}
echo = TRUE
mn_int[which.max(mn_int)]
```

3.Imputing missing values

```{r}
echo = TRUE
table(is.na(activity) == TRUE)
```

```{r}
summary(activity)
```

4. Filling in all of the Missing Values in the Dataset

```{r}
echo = TRUE
activity2 <- activity  # creation of the dataset that will have no more NAs
for (i in 1:nrow(activity)){
     if(is.na(activity$steps[i])){
         activity2$steps[i]<- mn_int[[as.character(activity[i, "interval"])]]
     }
}
```

Histogram of steps per day 

```{r}
echo = TRUE
su2 <- tapply(activity2$steps, activity2$date, sum, na.rm=T)
hist(su2, xlab = "sum of steps per day", main = "histogram of steps per day", col="green")
```

Remove Na's
```{r}
mean_su2 <- round(mean(su2))
median_su2 <- round(median(su2))
```

The Mean:

```{r}
echo = TRUE
print(c("The mean is",mean_su2))
```

The Median:

```{r}
print(c("The median is",median_su2))
```

Comparison of the new values with the old values

```{r}
echo = TRUE
df_summary <- rbind(df_summary, data.frame(mean = c(mean_su, mean_su2), median = c(median_su, median_su2)))
rownames(df_summary) <- c("with NA's", "without NA's")
print(df_summary)
```


```{r}
echo = TRUE
summary(activity2)
```

Confirmed there are no  more NA's in the steps variable.

4.Differences in Activity Patterns between Weekdays and Weekends:

New Colmun added to dataframe containing the factors weekday days & weekend days

```{r}
echo = TRUE
activity2$weekday <- c("weekday")
activity2[weekdays(as.Date(activity2[, 2])) %in% c("Saturday", "Sunday", "samedi", "dimanche", "saturday", "sunday", "Samedi", "Dimanche"), ][4] <- c("weekend")
table(activity2$weekday == "weekend")
```

```{r}
activity2$weekday <- factor(activity2$weekday)
```

 In order to visualize the difference between weekends and days of the week, a new dataframe is created to be usable by the lattice package.
 
```{r}
 echo = TRUE
activity2_weekend <- subset(activity2, activity2$weekday == "weekend")
activity2_weekday <- subset(activity2, activity2$weekday == "weekday")

mean_activity2_weekday <- tapply(activity2_weekday$steps, activity2_weekday$interval, mean)
mean_activity2_weekend <- tapply(activity2_weekend$steps, activity2_weekend$interval, mean)
```

```{r}
echo = TRUE
library(lattice)
df_weekday <- NULL
df_weekend <- NULL
df_final <- NULL
df_weekday <- data.frame(interval = unique(activity2_weekday$interval), avg = as.numeric(mean_activity2_weekday), day = rep("weekday", length(mean_activity2_weekday)))
df_weekend <- data.frame(interval = unique(activity2_weekend$interval), avg = as.numeric(mean_activity2_weekend), day = rep("weekend", length(mean_activity2_weekend)))
df_final <- rbind(df_weekday, df_weekend)

xyplot(avg ~ interval | day, data = df_final, layout = c(1, 2), 
       type = "l", ylab = "Number of steps")
```



